A Nodal Discontinuous Galerkin Method for the Multi-Group Two-Moment Model of Neutrino Transport

### 2018-11-13 Documents/M1/dgM1Curvilinear.tex
uncommitted changes

\newcommand{\ee}[1]{{\color{red} #1}}
\newcommand{\dx}{\Delta x}
\newcommand{\pbK}{\partial\mathbf{K}}
\newcommand{\UDG}{\bU_{\mbox{\tiny DG}}}
\newcommand{\sumx}{\sum_{i=1}^{d}}
\newcommand{\NES}{\mbox{\tiny NES}}

@@

\title{Nodal Discontinuous Galerkin Method for the Two-Moment Model of Neutrino Transport}
\author{Eirik Endeve\altaffilmark{1}, et al.}
\altaffiltext{1}{Computational and Applied Mathematics Group, Oak Ridge National Laboratory, Oak Ridge, TN 37831-6354, USA; endevee@ornl.gov}

\begin{abstract}
We present a nodal Discontinuous Galerkin (DG) method for solving the non-relativistic two-moment model for neutrino transport.  
The method --- ultimately targeted for simulations of core-collapse supernovae --- includes the use of curvilinear coordinates and non-ideal microphysics (nuclear equation of state and weak interactions) for neutrino-matter interactions.  
\end{abstract}

@@

In the following subsections we separately give the collision terms for the two-moment model by taking angular moments of Eq.~\eqref{eq:homogeneousBoltzmann}.  
(The total collision term in the two-moment model is then given by the sum of the separate terms.)  
The opacities are computed using the SFHo equation of state \citep{steiner_etal_2013}, and tabulated for densities $\rho\in[1.66\times10^{4},3.16\times10^{15}]$~g~cm$^{-3}$, temperatures $T\in[1.16\times10^{9},1.84\times10^{12}]$~K, and electron fractions $Y_{e}\in[0.01,0.6]$.  
The neutrino energies are in the range $\epsilonNu\in[0.1,300]$~MeV.  
We use logarithmic grids to cover the range in density $N_{\rho}=308$, temperature ($N_{T}=129$), and the neutrino energy ($N_{\epsilonNu}=40$), and a linear grid to cover the range in electron fraction ($N_{Y_{e}}=60$).  
The neutrino electron scattering rates are tabulated in terms of neutrino energies ($\epsilonNu$ and $\epsilonNu'$), temperature, and $\eta=\mu_{\nu}/(k_{\mbox{\tiny B}}T)$.  
Here we use a logarithmic grid in $\eta\in[1.0\times10^{-8},2.5\times10^{3}]$~MeV, using $N_{\eta}=100$ points.

\subsection{Emission and Absorption}

@@

\subsection{Elastic Scattering}

The isoenergetic scattering kernel $R_{\IS}(\bl'\cdot\bl,\epsilonNu)$ is taken as the sum of contributions from scattering on nucleons and nuclei
\begin{equation}
  R_{\IS}(\bl'\cdot\bl,\epsilonNu,\epsilonNu) = R_{\IS}^{\mbox{\tiny nucleon}}(\bl'\cdot\bl,\epsilonNu) + R_{\IS}^{\mbox{\tiny nuclei}}(\bl'\cdot\bl,\epsilonNu),
\end{equation}
where $R_{\IS}^{\mbox{\tiny nucleon}}$ and $R_{\IS}^{\mbox{\tiny nuclei}}$ are given by Eq. (35) and Eq. (36), respectively, in \citet{mezzacappaBruenn_1993a}.  
For inclusion in the two-moment model, these kernels are approximated by a two-term Legendre series
\begin{align}
  R_{\IS}(\cos\alpha,\epsilonNu) \approx \f{1}{2} \Phi_{\IS,0}(\epsilonNu) + \f{3}{2}\Phi_{\IS,1}(\epsilonNu)\,\cos\alpha.
\end{align}
By orthogonality of the Legendre polynomials ($P_{0}(x)=1,P_{1}(x)=x$), the expansion coefficients are given by
\begin{equation}
  \Phi_{\IS,\ell}(\epsilonNu)
  =\int_{-1}^{+1}R_{\IS}(\cos\alpha,\epsilonNu)\,P_{\ell}(\cos\alpha)\,d\cos\alpha \quad (\ell=0,1).  
\end{equation}
These integrals are performed numerically using a ??-point Gauss-Legendre quadrature rule.  (\ee{Ran: Fill in the number of quadrature points}.)
Inserting the Legendre expansion into Eq.~\eqref{eq:homogeneousBoltzmann} and taking the angular moments gives no contribution to the equation for $\cJ$ (energy conservation), while the contribution to the equation for $\bcH$ is
\begin{equation}
  \f{1}{c}\deriv{\bcH}{t}
  =-\big(\,\sigma_{\IS,0}-\sigma_{\IS,1}\big)\,\bcH,
\end{equation}
where we have defined
\begin{equation}
  \sigma_{\IS,\ell}(\epsilonNu)
  =\f{2\pi\,\epsilonNu^{2}}{c\,(hc)^{3}}\,\Phi_{\IS,\ell}(\epsilonNu) \quad (\ell=0,1).  
\end{equation}
We tabulate the scattering opacities $\sigma_{\IS,0}$ and $\sigma_{\IS,1}$, which have units cm$^{-1}$, but use only $\sigma_{\IS,0}$ in the numerical experiments in Section~\ref{sec:numericalTableOpacities}, where $\sigma=\sigma_{\IS,0}$ in Equation~\eqref{eq:momentEquationMomentum}.

@@

The NES scattering kernel $R_{\NES}^{\Out}(\cos\alpha,\epsilonNu,\epsilonNu')$ is given by Eq.~(9) in \citet{mezzacappaBruenn_1993c}.  
As we did in the isoenergetic scattering case, we approximate the angular dependence of the scattering kernels by a two-term Legendre expansion; e.g.,
\begin{equation}
  R^{\Out}_{\NES}(\cos\alpha,\epsilonNu,\epsilonNu') 
  \approx \f{1}{2} \Phi_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu') + \f{3}{2}\Phi^{\Out}_{\NES,1}(\epsilonNu,\epsilonNu')\,\cos\alpha,
\end{equation}
where
\begin{equation}
  \Phi_{\NES,\ell}^{\Out}(\epsilonNu,\epsilonNu')
  =\int_{-1}^{+1}R_{\NES}^{\Out}(\cos\alpha,\epsilonNu,\epsilonNu')\,P_{\ell}(\cos\alpha)\,d\cos\alpha \quad (\ell=0,1).  
\end{equation}
(Again, these integrals are computed numerically using a ??-point Gauss-Legendre quadrature.  (\ee{Ran: Fill in the number of quadrature points}.))
The in-scattering kernels $\Phi_{\NES,\ell}^{\In}(\epsilonNu,\epsilonNu')$ are obtained from symmetry properties \citep{cernohorsky_1994}
\begin{equation}
  \Phi_{\NES,\ell}^{\In}(\epsilonNu,\epsilonNu')=\Phi_{\NES,\ell}^{\Out}(\epsilonNu',\epsilonNu)
  \quad\text{and}\quad
  \Phi_{\NES,\ell}^{\Out}(\epsilonNu',\epsilonNu)
  =e^{-(\epsilonNu-\epsilonNu')/k_{\mbox{\tiny B}}T}\,\Phi_{\NES,\ell}^{\Out}(\epsilonNu,\epsilonNu').
  \label{eq:symmetryNES}
\end{equation}
Inserting the Legendre expansion into Eq.~\eqref{eq:homogeneousBoltzmann}, and taking the zeroth moment (number equation), gives
\begin{align}
  \f{1}{c}\deriv{\cJ}{t}
  &=\big(\,4\pi-\cJ(\epsilonNu)\,\big)\int_{\bbR^{+}}R_{\NES,0}^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  -\bcH(\epsilonNu)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\In}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'} \nonumber \\
  & \hspace{12pt}
  -\cJ(\epsilonNu)\int_{\bbR^{+}}R_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu')\,\big(\,4\pi-\cJ(\epsilonNu')\,\big)\,dV_{\epsilonNu'}
  +\bcH(\epsilonNu)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\Out}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'}.  
\end{align}
Similarly, taking the first moment (number flux) of Eq.~\eqref{eq:homogeneousBoltzmann} results in
\begin{align}
  \f{1}{c}\deriv{\bcH}{t}
  &=-\bcH(\epsilonNu)\int_{\bbR^{+}}R_{\NES,0}^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  +\big(\,(4\pi/3)\,\bI-\bcK\,\big)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\In}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'} \nonumber \\
  & \hspace{12pt}
  -\bcH\int_{\bbR^{+}}R_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu')\,\big(\,4\pi-\cJ(\epsilonNu')\,\big)\,dV_{\epsilonNu'}
  +\bcK\cdot\int_{\bbR^{+}}R_{\NES,1}^{\Out}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'},
\end{align}
where we have defined
\begin{equation}
  R_{\NES,0}^{\In/\Out}(\epsilonNu,\epsilonNu')
  =\f{1}{c\,(hc)^{3}}\,\f{1}{2}\,\Phi_{\NES,0}^{\In/\Out}(\epsilonNu,\epsilonNu')
  \quad\text{and}\quad
  R_{\NES,1}^{\In/\Out}(\epsilonNu,\epsilonNu')
  =\f{1}{c\,(hc)^{3}}\,\f{3}{2}\,\Phi_{\NES,1}^{\In/\Out}(\epsilonNu,\epsilonNu').  
\end{equation}
We tabulate the scattering kernels $R_{\NES,0}^{\Out}$ and $R_{\NES,1}^{\Out}$, which have units MeV$^{-3}$ cm$^{-1}$, and obtain the in-scattering rates from the symmetry properties in Eq.~\eqref{eq:symmetryNES}.  
We only use $R_{\NES,0}^{\In/\Out}$ in the numerical experiments in Section~\ref{sec:numericalTableOpacities}, so that $R^{\In/\Out}=R_{\NES,0}^{\In/\Out}$ in Eqs~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}.

@@

We benchmark the DG implementation against a series of test problems.  
We will vary the degree $k$ of the polynomial representation used in DG method from 0 to 3, which we then denote DG($k$).  
First we consider test problems in the streaming limit.  
In this case we use the explicit SSP-RK time integrators in Section~\ref{sec:explicitTime}.  
The 2-stage and 3-stage methods are denoted SSP-RK2 and SSP-RK3, respectively.  
As an example, an expected third-order accurate method (for problems with smooth solutions), using DG(2) and SSP-RK3, will then be denoted DG(2)+RK3.

@@

\subsection{Homogeneous Sphere}

The homogeneous sphere test \citep[cf.][]{smit_etal_1997} considers of a sphere with radius $R$.  
Inside the sphere ($r<R$), the absorption opacity and equilibrium distribution function are set to constant values $\chi=\chi_{0}$ and $f_{0}=1$, respectively.  
Outside the sphere, the absorption opacity is zero.  
The steady state solution, obtained by solving the transport equation in spherical symetry, is given by
\begin{equation}
  f_{\mbox{\tiny A}}(r,\mu)=f_{0}\,\big(1-e^{-\chi_{0}\,s(r,\mu)}\big),
  \label{distributionHomogeneousSphere}
\end{equation}
where
\begin{equation}
  s(r,\mu)
  =\left\{
  \begin{array}{lll}
    r\,\mu+R\,g(r,\mu) & \mbox{if}\quad r<R, & \mu\in[-1,+1], \\
    2\,R\,g(r,\mu) & \mbox{if}\quad r \ge R, & \mu\in[(1-(R/r)^{2})^{1/2},+1], \\
    0 & \mbox{otherwise},
  \end{array}
  \right.
\end{equation}
and $g(r,\mu)=[1-(r/R)^{2}(1-\mu^{2})]^{1/2}$.

@@

Results are plotted in Figure~\ref{fig:homogeneousSphere1D}.

@@

\section{Transport Problems with Tabulated Neutrino Opacities}
\label{sec:numericalTableOpacities}

For the tests in this section we employ the tabulated opacities described in Section~\ref{sec:opacities}.  
We present two tests from the previous section (the homogeneous sphere and spherical diffusion problems), where the background is kept fixed.  
Then we present a relaxation problem involving neutrino-electron scattering (NES) with a homogeneous background.  
Finally, we present a deleptonization problem where the background density mimics that of a collapsed stellar core.  
The first deleptonization problem (Em/Ab) involves emission and absorption only, while the second deleptonization problem (Full) involves the full set of opacities described in Section~\ref{sec:opacities}.

@@

\subsubsection{Emission and Absorption}

>>>>

\newcommand{\ee}[1]{{\color{red} #1}}
\newcommand{\dx}{\Delta x}
\newcommand{\pbK}{\partial\mathbf{K}}
\newcommand{\UDG}{\bU_{\mbox{\tiny DG}}}
\newcommand{\sumx}{\sum_{i=1}^{d}}
\newcommand{\NES}{\mbox{\tiny NES}}
\newcommand{\thornado}{\texttt{thornado}}

@@

\title{A Nodal Discontinuous Galerkin Method for the Multi-Group Two-Moment Model of Neutrino Transport}
\author{Eirik Endeve\altaffilmark{1}, et al.}
\altaffiltext{1}{Computational and Applied Mathematics Group, Oak Ridge National Laboratory, Oak Ridge, TN 37831-6354, USA; endevee@ornl.gov}

\begin{abstract}
We present a nodal Discontinuous Galerkin (DG) method for solving the non-relativistic multi-group two-moment model for neutrino transport.  
The method --- ultimately targeted for simulations of core-collapse supernovae --- includes the use of curvilinear coordinates and non-ideal microphysics (nuclear equation of state and weak interactions) for neutrino-matter interactions.  We give implementation details and present results from an extensive set of benchmark problems in one spatial dimension. 
\end{abstract}

@@

In the following subsections we separately give the collision terms for the two-moment model by taking angular moments of Eq.~\eqref{eq:homogeneousBoltzmann}.  
(The total collision term in the two-moment model is then given by the sum of the separate terms.)  
The opacities are computed using the SFHo equation of state \citep{steiner_etal_2013}, and tabulated for densities $\rho\in[1.66\times10^{4},3.16\times10^{15}]$~g~cm$^{-3}$, temperatures $T\in[1.16\times10^{9},1.84\times10^{12}]$~K, and electron fractions $Y_{e}\in[0.01,0.6]$.  
The neutrino energies are in the range $\epsilonNu\in[0.1,300]$~MeV.  
We use logarithmic grids to cover the range in density $N_{\rho}=308$, temperature ($N_{T}=161$), and the neutrino energy ($N_{\epsilonNu}=40$), and a linear grid to cover the range in electron fraction ($N_{Y_{e}}=60$).  
The neutrino electron scattering rates are tabulated in terms of neutrino energies ($\epsilonNu$ and $\epsilonNu'$), temperature, and $\eta=\mu_{\nu}/(k_{\mbox{\tiny B}}T)$.  
Here we use a logarithmic grid in $\eta\in[1.0\times10^{-8},2.5\times10^{3}]$~MeV, using $N_{\eta}=100$ points.

\subsection{Emission and Absorption}
\label{sec:electronCapture}

@@

\subsection{Elastic Scattering}
\label{sec:elasticScattering}

The isoenergetic scattering kernel $R_{\IS}(\bl'\cdot\bl,\epsilonNu)$ is taken as the sum of contributions from scattering on nucleons and nuclei
\begin{equation}
  R_{\IS}(\bl'\cdot\bl,\epsilonNu,\epsilonNu) = R_{\IS}^{\mbox{\tiny nucleon}}(\bl'\cdot\bl,\epsilonNu) + R_{\IS}^{\mbox{\tiny nuclei}}(\bl'\cdot\bl,\epsilonNu),
\end{equation}
where $R_{\IS}^{\mbox{\tiny nucleon}}$ and $R_{\IS}^{\mbox{\tiny nuclei}}$ are given by Eq. (35) and Eq. (36), respectively, in \citet{mezzacappaBruenn_1993a}.  
For inclusion in the two-moment model, these kernels are approximated by a two-term Legendre series
\begin{align}
  R_{\IS}(\cos\alpha,\epsilonNu) \approx \f{1}{2} \Phi_{\IS,0}(\epsilonNu) + \f{3}{2}\Phi_{\IS,1}(\epsilonNu)\,\cos\alpha.
\end{align}
By orthogonality of the Legendre polynomials ($P_{0}(x)=1,P_{1}(x)=x$), the expansion coefficients are given by
\begin{equation}
  \Phi_{\IS,\ell}(\epsilonNu)
  =\int_{-1}^{+1}R_{\IS}(\cos\alpha,\epsilonNu)\,P_{\ell}(\cos\alpha)\,d\cos\alpha \quad (\ell=0,1).  
\end{equation}
These integrals are performed numerically using a 20-point Gauss-Legendre quadrature rule.  
Inserting the Legendre expansion into Eq.~\eqref{eq:homogeneousBoltzmann} and taking the angular moments gives no contribution to the equation for $\cJ$ (energy conservation), while the contribution to the equation for $\bcH$ is
\begin{equation}
  \f{1}{c}\deriv{\bcH}{t}
  =-\big(\,\sigma_{\IS,0}-\sigma_{\IS,1}\big)\,\bcH,
\end{equation}
where we have defined
\begin{equation}
  \sigma_{\IS,\ell}(\epsilonNu)
  =\f{2\pi\,\epsilonNu^{2}}{c\,(hc)^{3}}\,\Phi_{\IS,\ell}(\epsilonNu) \quad (\ell=0,1).  
\end{equation}
We tabulate the scattering opacities $\sigma_{\IS,0}$ and $\sigma_{\IS,1}$, which have units cm$^{-1}$, but use only $\sigma_{\IS,0}$ in the numerical experiments in Section~\ref{sec:numericalTableOpacities}, where $\sigma=\sigma_{\IS,0}$ in Equation~\eqref{eq:momentEquationMomentum}.

@@

The NES scattering kernel $R_{\NES}^{\Out}(\cos\alpha,\epsilonNu,\epsilonNu')$ is given by Eq.~(9) in \citet{mezzacappaBruenn_1993c}.  
As we did in the isoenergetic scattering case, we approximate the angular dependence of the scattering kernels by a two-term Legendre expansion; e.g.,
\begin{equation}
  R^{\Out}_{\NES}(\cos\alpha,\epsilonNu,\epsilonNu') 
  \approx \f{1}{2} \Phi_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu') + \f{3}{2}\Phi^{\Out}_{\NES,1}(\epsilonNu,\epsilonNu')\,\cos\alpha,
\end{equation}
where
\begin{equation}
  \Phi_{\NES,\ell}^{\Out}(\epsilonNu,\epsilonNu')
  =\int_{-1}^{+1}R_{\NES}^{\Out}(\cos\alpha,\epsilonNu,\epsilonNu')\,P_{\ell}(\cos\alpha)\,d\cos\alpha \quad (\ell=0,1).  
\end{equation}
(These integrals are computed numerically using a 30-point Gauss-Legendre quadrature.  
The in-scattering kernels $\Phi_{\NES,\ell}^{\In}(\epsilonNu,\epsilonNu')$ are obtained from symmetry properties \citep{cernohorsky_1994}
\begin{equation}
  \Phi_{\NES,\ell}^{\In}(\epsilonNu,\epsilonNu')=\Phi_{\NES,\ell}^{\Out}(\epsilonNu',\epsilonNu)
  \quad\text{and}\quad
  \Phi_{\NES,\ell}^{\Out}(\epsilonNu',\epsilonNu)
  =e^{-(\epsilonNu-\epsilonNu')/k_{\mbox{\tiny B}}T}\,\Phi_{\NES,\ell}^{\Out}(\epsilonNu,\epsilonNu').
  \label{eq:symmetryNES}
\end{equation}
Inserting the Legendre expansion into Eq.~\eqref{eq:homogeneousBoltzmann}, and taking the zeroth moment (number equation), gives
\begin{align}
  \f{1}{c}\deriv{\cJ}{t}
  &=\big(\,4\pi-\cJ(\epsilonNu)\,\big)\int_{\bbR^{+}}R_{\NES,0}^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  -\bcH(\epsilonNu)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\In}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'} \nonumber \\
  & \hspace{12pt}
  -\cJ(\epsilonNu)\int_{\bbR^{+}}R_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu')\,\big(\,4\pi-\cJ(\epsilonNu')\,\big)\,dV_{\epsilonNu'}
  +\bcH(\epsilonNu)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\Out}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'}.  
\end{align}
Similarly, taking the first moment (number flux) of Eq.~\eqref{eq:homogeneousBoltzmann} results in
\begin{align}
  \f{1}{c}\deriv{\bcH}{t}
  &=-\bcH(\epsilonNu)\int_{\bbR^{+}}R_{\NES,0}^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  +\big(\,(4\pi/3)\,\bI-\bcK\,\big)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\In}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'} \nonumber \\
  & \hspace{12pt}
  -\bcH\int_{\bbR^{+}}R_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu')\,\big(\,4\pi-\cJ(\epsilonNu')\,\big)\,dV_{\epsilonNu'}
  +\bcK\cdot\int_{\bbR^{+}}R_{\NES,1}^{\Out}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'},
\end{align}
where we have defined
\begin{equation}
  R_{\NES,0}^{\In/\Out}(\epsilonNu,\epsilonNu')
  =\f{1}{c\,(hc)^{3}}\,\f{1}{2}\,\Phi_{\NES,0}^{\In/\Out}(\epsilonNu,\epsilonNu')
  \quad\text{and}\quad
  R_{\NES,1}^{\In/\Out}(\epsilonNu,\epsilonNu')
  =\f{1}{c\,(hc)^{3}}\,\f{3}{2}\,\Phi_{\NES,1}^{\In/\Out}(\epsilonNu,\epsilonNu').  
\end{equation}
We tabulate the scattering kernels $R_{\NES,0}^{\Out}$ and $R_{\NES,1}^{\Out}$, which have units MeV$^{-3}$ cm$^{-1}$, and obtain the in-scattering rates from the symmetry properties in Eq.~\eqref{eq:symmetryNES}.  
We only use $R_{\NES,0}^{\In/\Out}$ in the numerical experiments in Section~\ref{sec:numericalTableOpacities}, so that $R^{\In/\Out}=R_{\NES,0}^{\In/\Out}$ in Eqs~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}.

@@

We benchmark the DG implementation against a series of test problems.  
We will vary the degree $k$ of the polynomial representation used in DG method from 0 to 3, which we then denote DG($k$).  
First we consider test problems in the streaming limit.  
In this case we use the explicit SSP-RK time integrators in Section~\ref{sec:explicitTime}.  
The 2-stage and 3-stage methods are denoted SSP-RK2 and SSP-RK3, respectively.  
As an example, an expected third-order accurate method (for problems with smooth solutions), using DG(2) and SSP-RK3, will then be denoted DG(2)+RK3.  
For tests involving collisions we use the two-stage Semi-Implicit Runge-Kutta method described in Section~\ref{sec:semiImplicitTime}.  
When this time integrator is combined with DG($k$) spatial discretization, the method is denoted DG($k$)+SIRK2.

@@

\subsection{Homogeneous Sphere}
\label{sec:homogeneousSphere}

The homogeneous sphere test \citep[cf.][]{smit_etal_1997} considers of a sphere with radius $R$.  
Inside the sphere ($r<R$), the absorption opacity and equilibrium distribution function are set to constant values $\chi=\chi_{0}$ and $f_{0}=1$, respectively.  
Outside the sphere, the absorption opacity is zero.  
The steady state solution, obtained by solving the transport equation in spherical symmetry, is given by
\begin{equation}
  f_{\mbox{\tiny A}}(r,\mu)=f_{0}\,\big(1-e^{-\chi_{0}\,s(r,\mu)}\big),
  \label{distributionHomogeneousSphere}
\end{equation}
where
\begin{equation}
  s(r,\mu)
  =\left\{
  \begin{array}{lll}
    r\,\mu+R\,g(r,\mu) & \mbox{if}\quad r<R, & \mu\in[-1,+1], \\
    2\,R\,g(r,\mu) & \mbox{if}\quad r \ge R, & \mu\in[(1-(R/r)^{2})^{1/2},+1], \\
    0 & \mbox{otherwise},
  \end{array}
  \right.
\end{equation}
and $g(r,\mu)=[1-(r/R)^{2}(1-\mu^{2})]^{1/2}$.

@@

Results are plotted in Figure~\ref{fig:homogeneousSphere1D}.  
In the left panels we plots results obtained for the thick case, while results for the thin case are plotted in the right panels.  
We plot the radiation density $\cJ$ in the top panels, the flux factor $\cH_{r}/\cJ$ in the middle panels, and the Eddington factor $\psi$ in the bottom panels.  
The two-moment model captures the analytic solution reasonably well in both the thin and the thick cases, especially for the density $\cJ$.  
The discrepancy in the flux and Eddington factors between the numerical results and the analytical solution is comparable to that of other implementations of the two-moment model \citep[e.g.,][]{smit_etal_1997}.

@@

\section{Multi-group Transport Problems with Tabulated Neutrino Opacities}
\label{sec:numericalTableOpacities}

For the tests in this section we employ some of the tabulated opacities described in Section~\ref{sec:opacities}.  
In Section~\ref{sec:homogeneousSphereTable}, we present a multi-group version of the homogeneous sphere test from Section~\ref{sec:homogeneousSphere}, where the background is kept fixed.  
In Section~\ref{sec:deleptonizationProblem}, we present a deleptonization problem where the initial background density mimics that of a collapsed stellar core.  
The deleptonization problem uses the tabulated electron capture and elastic scattering rates on nucleons and nuclei as described in Sections~\ref{sec:electronCapture} and \ref{sec:elasticScattering}.  
For this problem, the temperature and electron fraction of the background are evolved with the radiation field as described in Section~\ref{sec:couplingElectronCapture} below.  
The deleptonization problem is distributed with the mini-app, available at \url{https://github.com/ECP-Astro/thornado_mini}.

@@

%\subsubsection{Emission and Absorption}
\label{sec:couplingElectronCapture}

Electron capture on nucleons and nuclei results in changes to the electron fraction and internal energy of the fluid given by
\begin{align}
  d_{t}Y_{e}
  &=-\f{m_{\mbox{\tiny B}}}{\rho}\int_{\bbR^{+}}\tilde{\chi}\,\big(\cJ_{0}-\cJ\big)\,dV_{\epsilonNu}, \label{eq:electronFractionEmAb} \\
  d_{t}\epsilon
  &=-\f{1}{\rho}\int_{\bbR^{+}}\tilde{\chi}\,\big(\cJ_{0}-\cJ\big)\,\epsilonNu\,dV_{\epsilonNu}, \label{eq:internalEnergyEmAb}
\end{align}
where $m_{\mbox{\tiny B}}$ is the baryon mass and $\epsilon$ is the specific internal energy.  
In each implicit step in the SIRK2 scheme discussed in Section~\ref{sec:semiImplicitTime} we solve Eqs.~\eqref{eq:electronFractionEmAb}-\eqref{eq:internalEnergyEmAb} coupled to the collision part of the moment equations.  
\begin{align}
  Y_{e}^{n+1}-Y_{e}^{n}
  &=-\f{m_{\mbox{\tiny B}}}{\rho}\sum_{j=1}^{M^{\epsilonNu}}W_{j}^{(2)}\,\gamma_{j}^{n}\,\big(\cJ_{0,j}^{n+1}-\cJ_{j}^{n+1}\big), \label{eq:electronFractionEmAbDiscrete0} \\
  \epsilon^{n+1}-\epsilon^{n}
  &=-\f{1}{\rho}\sum_{j=1}^{M^{\epsilonNu}}W_{j}^{(3)}\,\gamma_{j}^{n}\,\big(\cJ_{0,j}^{n+1}-\cJ_{j}^{n+1}\big), \label{eq:internalEnergyEmAbDiscrete0}
\end{align}
where $W_{j}^{(2,3)}=w_{q}\,\epsilonNu_{i,q}^{2,3}\,\Delta\epsilonNu_{i}$ and $\gamma_{j}^{n}=\dt\,\tilde{\chi}_{j}^{n}$.  
Note that the absorption opacity is defined at the known time level $t^{n}$, which avoids computation of derivatives with respect to temperature and electron fraction.  
The radiation energy density is updated from
\begin{equation}
  \cJ_{j}^{n+1}=\f{\cJ_{j}^{n}+\gamma_{j}^{n}\,\cJ_{0,j}^{n+1}}{1+\gamma_{j}^{n}}, 
  \label{eq:momentEquationEnergyEmAbDiscrete}
\end{equation}
which, when inserted into Eqs.~\eqref{eq:electronFractionEmAbDiscrete0}-\eqref{eq:internalEnergyEmAbDiscrete0}, results in a nonlinear system for the unknowns $Y^{n+1}$ and $\epsilon^{n+1}$
\begin{align}
  Y_{e}^{n+1}-Y_{e}^{n}
  &=-\f{m_{\mbox{\tiny B}}}{\rho}\sum_{j=1}^{M^{\epsilonNu}}W_{j}^{(2)}\,\tilde{\gamma}_{j}^{n}\,\big(\cJ_{0,j}^{n+1}-\cJ_{j}^{n}\big), \label{eq:electronFractionEmAbDiscrete} \\
  \epsilon^{n+1}-\epsilon^{n}
  &=-\f{1}{\rho}\sum_{j=1}^{M^{\epsilonNu}}W_{j}^{(3)}\,\tilde{\gamma}_{j}^{n}\,\big(\cJ_{0,j}^{n+1}-\cJ_{j}^{n}\big), \label{eq:internalEnergyEmAbDiscrete}
\end{align}
where $\tilde{\gamma}_{j}^{n}=\gamma_{j}^{n}/(1+\gamma_{j}^{n})$.  
Since the equilibrium density $\cJ_{0,j}$ is a nonlinear function of $Y_{e}$ and $\epsilon$, Eqs.~\eqref{eq:electronFractionEmAbDiscrete}-\eqref{eq:internalEnergyEmAbDiscrete} are solved using Newton's method.  
Once $Y_{e}^{n+1}$ and $\epsilon^{n+1}$ are obtained, the radiation density is obtained from Eq.~\eqref{eq:momentEquationEnergyEmAbDiscrete}, while the flux is given by
\begin{equation}
  \bcH_{j}^{n+1}=\bcH_{j}^{n}/(1+\gamma_{j}^{n}).  
  \label{eq:momentEquationMomentumEmAbDiscrete}
\end{equation}
When elastic scattering is included, the above description remains unchanged, except for $\gamma_{j}^{n}\to\dt\big(\tilde{\chi}_{j}^{n}+\sigma_{j}^{n}\big)$ in Eq.~\eqref{eq:momentEquationMomentumEmAbDiscrete}.
